---
title: 'Statistical report: Differences in HF phenotype between incident HF patients with and without prior RA'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.1.7. 

## Inclusion/exclusion criteria

```{r flow}
mykable(flow, caption = "Flowchart")
mykable(ncontrols, caption = "Number of controls / case", scale_down = F)
```

A period of 14 days is allowed between first registration in NPR and registration in SwedeHF.

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: `r max(pdata$shf_indexdtm)`.  

## HF/EF

By excluding cases with previous diagnosis of HF in NPR/SwedeHF the aim is to keep only incident HF cases and thereby
aim to discuss directionality. 

EF is recorded in SwedeHF. The primary analysis will be on EF </=> 40% due to the small number of patients 
with previous RA. Analyses EF <30% (rEF), 40-49% (mrEF) and =>50% (pEF) will also be presented.  

## In/out-patients

The main analysis will be on patients (and their controls) recorded as out-patients in SwedeHF. 
In-patients will be presented separately. 

## Controls for HF

Eligible controls were persons without re-used and changed Swedish personal identity numbers (PINs) and free of HF, defined either in the National Patient registry (ICD-10: I110, I130, I132, I255, I420, I423, 1425-9, I43, I50, J81, K761, R57, ICD-9: 414W, 425E-H, 425W, 425X, 428) or SwedeHF, at the end of the year of the index date for the case and matched by gender, year of birth and county of residence at index. For posts with age >= 90 years, controls were only matched by gender and year of birth due to the smaller available number of controls. The selection of controls was performed without replacement.

## Definition of RA 

Two visits (spec öppenvård) with HDIA M05, M06, where at least on of the visits 
should be at an internal medicin/reumatological ward (MVO 101, 131). 
The onset of RA diagnosis is then defined to the second visit. If a visit with 
HDIA M073, L405, M45, M46, M32 occurs prior to the second visit this patient 
is not defined as a RA. 

### Blankning

No blanking is applied in the main analysis. As a sensitivity analysis a blanking period of 6 months is applied. 

## Variables created specifically for this project

```{r sos}
mykable(metaout, caption = "Additional outcomes from NPR", scale_down = F)
```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

Vill ni ha tester här? Mellan vad? Ska ra vara med i tabell? 

## Association between incident HF and prior RA

The association between incident HF and prior RA is investigated 
using a conditional logistic regression where the outcome is if a person 
is a HF case or a control. The matched "pairs" are included as a strata variable and 
EF and RA as main effects and with an interaction term. 
In this analysis both cases and controls are included. 

Adjustment is performed for variables that are present at the 
time of the incident HF, and for the controls, the matched case's time. 

Adjustment is partly performed for `r paste0(modvars1, collapse = ", ")` 

and partly `r paste0(modvars2, collapse = ", ")`.

Number of available observations for each group: 
`r mykable(pdata %>% count(shf_location, casecontrol, sos_com_ra, shf_ef_cat), scale_down = FALSE)` 

\clearpage
\newpage

```{r, child = "../src/cortab.Rmd"}

```

\clearpage
\newpage

## Outcomes after HF

For this analysis only cases (patients with HF) are selected. 
The patients with prior RA are compared to those without prior RA. 

The following outcomes are considered: 

- Cardiovascular mortality/first hospitalization for HF/stroke/MI (primary endpoint)
- First hospitalization for HF
- All-cause mortality

Incidence per 1000 py is calculated with 95% Poisson confidence intervals. 
The outcomes are presented with 1-KM curves. 
Cox proportional hazards regressions are used to evaluate the association 
between prior RA and the respective outcomes. The interaction between 
prior RA and EF is included in the model and adjustment performed on variables that are present at the 
time of the incident HF for the same variables as in the previous models. 
Data are censored at 2018-12-31 or death/emigration. 

\clearpage
\newpage

```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/coxtab.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/rahf. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
